---
title: "Analise Weg"
author: "Carlos E. Carvalho"
date: "1/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading to files about the company WEG:
 
 3t20_dre.csv: file with the company results #https://ri.weg.net/informacoes-financeiras/planilhas/
 
 WEGE3SA.csv: file with stock market prices and volume
 https://br.financas.yahoo.com/quote/WEGE3.SA/history?period1=1262217600&period2=1611014400&interval=1d&filter=history&frequency=1d&includeAdjustedClose=true
 
 I am trying to see how the stock prices react when the company results are presented.
 Everything I did here, I've learned in the course Formação Cientista de Dados, at DataScience Academy
 https://www.datascienceacademy.com.br/bundles?bundle_id=formacao-cientista-de-dados
 
 Carlos Eduardo Carvalho
 carlos.e.carvalho@gmail.com
 Linkedin: https://www.linkedin.com/in/carlos-carvalho-93204b13/
 https://github.com/CarlosCarvalho1981/Analise_Weg.git

```{r}
setwd("D:/CIENTISTA_DADOS/WEG")
getwd()
```

Loading packages 

```{r echo = FALSE}
library(dplyr)
library(readr)
library(data.table)
library(ggplot2)
library(reshape2)
library(plotrix)
```

Loading the file
```{r}
dre <- read_csv("3t20_dre.csv")
dre
```
There are some problems with the way the rows and columns are.
Let's see the class of the object
```{r}
class(dre)
```
Let's remove the first column that won't help us in any way.
```{r}
dre$X1 <- NULL
```
Let's swap the rows and the columns.
```{r}
inv_dre <- t(dre)

```
Now it became a matrix
```{r}
class(inv_dre)
```
Let's turn it into a data frame
```{r}
inv_dre <- as.data.frame(inv_dre)
class(inv_dre)
```
Since the name of some columns have characters that aren't recognized, let's rename everything.
```{r}
colnames(inv_dre) <- c("Data", "Receita de Venda de Bens e/ou Servicos", "Custo dos Bens e/ou Servicos Vendidos", 
                       "Resultado Bruto" , "Despesas/Receitas Operacionais", "Despesas com Vendas", "Despesas Gerais e Administrativas" ,
                       "Honorarios dos Administradores", "Outras Despesas Administrativas", "Outras Receitas Operacionais", "Outras Despesas Operacionais",
                       "Resultado de Equivalencia Patrimonial", "Resultado Antes do Resultado Financeiro e dos Tributos",
                       "Resultado Financeiro", "Receitas Financeiras", "Despesas Financeiras", "Resultado Antes dos Tributos sobre o Lucro",
                       "Imposto de Renda e Contribuicao Social sobre o Lucro", "Corrente", "Diferido", "Resultado Liquido das Operacoes Continuadas",
                       "Lucro/Prejuizo Consolidado por Periodo", "Atribuido a Socios da Empresa Controladora", "Atribuido a Socios nao Controladores",
                       "Lucro por Acao - (Reais/Acao", "Lucro Basico por Acao", "ON", "Lucro Diluido por Acao", "ON")

```
We still have a row that is not usefull.  Let's remove it
```{r}
inv_dre <- inv_dre[-c(1),]
```
And also remove the last 2 columns
```{r}
inv_dre <- inv_dre[,-c(30,31)]
```
Now we have a good data frame!
Let's take the columns Data and Lucro/Prejuizo Consolidado por Periodo.
```{r}
sub_dre <- inv_dre[,c("Data", "Lucro/Prejuizo Consolidado por Periodo")]
```
How does this subset looks like?
```{r}
head(sub_dre)
```
Let's see the types of the columns
```{r}
glimpse(sub_dre)
```
Both variables are factors.  We're going to need the variable Data as a date format.
Let's extract a vector to make the conversion
```{r}
sub_dre1 <- sub_dre$Data
Data2 <- as.Date(sub_dre1, format = "%m/%d/%Y")
glimpse(sub_dre1)
```
And we put it back together
```{r}
sub_dre <- cbind(sub_dre, Data2)
head(sub_dre)
```
We have now columns with the same information.  Let's take off the one we don't need anymore.
```{r}
sub_dre <- sub_dre[-c(1)]
```
In order to make a plot, let's take only the last 10 rows.
```{r}
sub_dre2 <- sub_dre[34:44,]
sub_dre2
```
We are going to see a scatter plot where the colors of the points are related to the column `Lucro/Prejuizo Consolidado por Periodo`
```{r}
ggplot(data = sub_dre2, aes(x=Data2, y = `Lucro/Prejuizo Consolidado por Periodo`, colour = `Lucro/Prejuizo Consolidado por Periodo` ))+
  geom_point(aes(size = 2)) + xlab("Data") + ylab("Lucro/Prejuizo por Periodo")

```
Now we're going to load the file with WEG stock prices
```{r}
acoes <- read_csv("WEGE3SA.csv", col_types = list(
  Date = col_date(),
  Open = col_double(),
  High = col_double(),
  Low = col_double(),
  Close = col_double(),
  'Adj Close' = col_double(),
  Volume = col_integer()
))
```
The file has some NULL rows that's why we've got those warnings.
Let's see.
```{r}
acoes
```
The NULL rows became NA.
Let's drop those NA rows
```{r}
acoes <- acoes[complete.cases(acoes), ]
```
Let's get a subset with the last 30 rows, just to see...
```{r}
acoes30 <- acoes[661:691,]
```
#The last 2 are NA to.
```{r}
acoes30 <- acoes30[complete.cases(acoes30), ]
acoes30
```
#We are going to see a scatter plot of the stock prices with a nice linear regression.
```{r}
ggplot(data = acoes30, aes(x = Date, y = Close))+
  geom_point() + xlab("Data") + ylab("Valor de Fechamento") + 
  geom_smooth(method = lm, color = "blue", se = FALSE)
```
We can see that there's a tendency of growing on the prices
Now we're going to extract some vectors to work with.
We are going to take the dates that are in both data frames: sub_dre2 and acoes. Or, at least, the next dates. Had to do that manually.
```{r}
acoes2 <- acoes[c(1,63,126,185,244,306,371,432,493,554,619),] 
acoes2
Close2 <- acoes2$Close
Volume2 <- acoes2$Volume
Close2
```
Let's have a column with the range of values instead of the real values.
First, looking for the max and min values of the stock prices.
```{r}
maxClose <- max(acoes$Close, na.rm = TRUE)
maxClose 
minClose <- min(acoes$Close, na.rm = TRUE)
minClose 


groupClose <- function(fechamento){
  if(fechamento > 15.00 & fechamento <= 30.00){
    return(30)
  }else if(fechamento > 30.00 & fechamento <= 45.00){
    return(45)
  }else if(fechamento > 45.00 & fechamento <= 60.00){
    return(60)
  }else if(fechamento > 60.00 & fechamento <= 75.00){
    return(75)
  }else if(fechamento > 75.00 & fechamento <= 90.00){
    return(90)
  }else{
    return(95)
  }
}

acoes2$CloseGroup <- sapply(acoes2$Close, groupClose)
View(acoes2)

CloseGroup2 <- acoes2$CloseGroup
```
And now we insert the vectors into the data frame
```{r}
sub_dre2 <- cbind(sub_dre2, Close2)
sub_dre2 <- cbind(sub_dre2, Volume2)
sub_dre2 <- cbind(sub_dre2, CloseGroup2)
colnames(sub_dre2) <- c("Lucro/Prejuizo", "Data2", "Close2", "Volume2", "CloseGroup2")
head(sub_dre2)
```
Let's see a scatter plot of  Lucro/Prejuizo Consolidado por Periodo, with the size of the dots as the
closing stock prices and the color as the volume.
```{r}
ggplot(data = sub_dre2, aes(x=Data2, y = `Lucro/Prejuizo`))+
  geom_point(aes(size = CloseGroup2, color = Volume2)) + xlab("Data") + ylab("Lucro/Prejuizo por Periodo")

```
Now we are going to plot two line plots: Lucro/Prejuizo and Preco de Fechamento. We want to see if the behavior of the stock price is the same as the behavior of the results of the company.
```{r}
par(mar = c(4,3,3,3), col.axis = "black", adj = 0.5)
plot(sub_dre2$Data2, sub_dre2$`Lucro/Prejuizo`, type = "l", col = "red", bty = "o", 
     xlab = "Data", ylab = "Lucro/Prejuizo")
par(new = T)

plot(sub_dre2$Data2, sub_dre2$Close2, type = "l", ann = F, axes = F, col = "darkblue", ylab = "Preco")
axis(side = 4)
legend("bottomright", pch = 1, col = c("blue", "red"), legend = c("Preco de Fechamento", "Lucro/Prejuizo"))
par(mfrow=c(1,1))
title(main = "Preco x Lucro/Prejuizo")
```
Looks like it is !
Now let's look at the volume of the company stocks during the days the results of the company were presented.
```{r}
ggplot(data = sub_dre2, aes(x = Data2, y = Volume2 )) +
  geom_col(aes(fill = Volume2)) +
  ylab("Volume de Ações Negociadas") + xlab("Data de Divulgação dos Resultados")
```
Now let's create another dataset, with dates next to the date when the results of the company were given.
```{r}
acoes3 <- acoes[c(1,2,3,62,63,64,125,126,127,184,185,186,243,244,245,305,306,307,370,371,372,431,432,433,492,493,494,553,554,555,618619,620),]
```
Dropping the NA rows.
```{r}
acoes3 <- acoes3[complete.cases(acoes3), ]
View(acoes3)
```
Creating the last plot to see the stock prices near the days when the results came:
```{r}
plot(acoes3$Date, acoes3$Close, type = "l", col = "red", bty = "o", 
     xlab = "Data", ylab = "Preço de fechamento")
```

